\bf On Devon Allen's Disqualification at the 2022 World Track and Field Championships

### 2025-02-27 Manuscript/manuscript.tex
responded to Prof Yan's comments

\jy{State the hypotheses here before diving into data/methods/results.}

@@

\jy{This two paragraphs are repetitive. Condense.}
To investigate whether RTs at the 2022 World Championships
were significantly different from other competitions, we
used data from male athletes in the 110-meter hurdles and 100-meter
dash, and female athletes in the 100-meter hurdles and 100-meter dash,
provided they competed in the 2022 World Championships and at least
one other competition (2022 national championships, 2019 World
Championships, or 2023 World Championships).  We excluded data from
200-meter dashes and longer events, as their RTs were found
to be significantly different. Negative RTs were excluded from the analysis,
while positive disqualified RTs were included due to their low frequency and not
being obvious outliers.

This study focuses on RTs from sprint events where reaction
time plays a critical role. Specifically, we included data from men’s
100-meter dashes and 110-meter hurdles, as well as women’s 100-meter
dashes and 100-meter hurdles. Statistical tests showed no significant
differences in RTs between these events, supporting their
inclusion in a unified analysis. However, we excluded data from
200-meter dashes and longer events, as their RTs were found
to be significantly different. This difference is not unexpected, as
RTs are generally less critical in longer sprint events
compared to shorter distances where explosive starts are paramount.
Negative RTs were excluded from the analysis, while positive
disqualified RTs were included due to their low frequency and not
being obvious outliers. % A
% sensitivity analysis regarding the inclusion of positive disqualified
% RTs is provided in the Supplementary Material.

@@

\jy{Opening sentence not accurate. This is the data to test the first
  hypothesis: national versus World Championship 2022.}

To assess the abnormality of the 2022 World Championships, we examined
data beyond World Championship events. This decision was motivated by
an exploratory analysis of how United States (US) athletes performed
at the 2022 US Track and Field Championships, held from June 23--26,
2022, at Hayward Field in Eugene, Oregon. Since this venue also hosted
the 2022 World Championships in August, it
provided a unique opportunity to compare. There were four US
110-meter hurdle athletes—Trey Cunningham, Daniel Roberts,
Grant Holloway, and Devon Allen—who competed in both events. All four
athletes recorded faster RTs in every World Championships race
compared to their performances at the national-level event. Similarly,
all four US 100-meter dash athletes---Marvin Bracy, Fred Kerley,
Travyon Bromwell, and Christian Coleman---also recorded slower RTs at
the US Championships than at the World Championships.
To expand the dataset, we included RTs from 100-meter hurdle
and 100-meter dash athletes who competed in other national
competitions held between May and July 2022 across various
countries. Compiling this data presented challenges, as RTs were not
centrally archived and often required searching country-specific
websites, with many results recorded in native languages.

@@

\jy{Improve the topic sentence: this data is for the second
  hypothesis: 2019 vs 2022; 2023 vs 2022.}
To compare RTs at the 2022 World Championships with those
recorded in the 2019 and 2023 World Championships, we prepared datasets
including RTs from athletes who competed in the 2022 World
Championships and at least one of the 2019 or 2023 World Championships.
Specifically, we identified athletes who participated in both the 2019
and 2022 World Championships, as well as those who participated in both
the 2022 and 2023 World Championships. In the 2019-2022 comparison,
RTs from 2022 were treated as the `treatment' group, with
2019 serving as the `control' group. Similarly, in the 2022-2023
comparison, RTs from 2022 were treated as the `treatment'
group, with 2023 serving as the `control' group. This structure allowed
us to prepare datasets suitable for examining RTs of athletes
who competed across multiple World Championships.

@@

\jy{write a quick overview paragraph here.}

>>>>

\jy{State the hypotheses here before diving into data/methods/results.}
\of{Like this?  I feel like the first paragraph of the data section serves as
a good introductino to this section. I dont' know what to write without
re-writing the introduction or what is discussed in the first paragraph of 2.1}

This investigation began to determine whether athletes who competed at the 2022
Championships had significantly different RTs to when those same athletes
competed in comparable and relevant events.

\of{or like this?}

Within our investigation to determine if athletes performed abnormally in 2022,
we will analysze times from athletes who competed in earlier meets in 2022 
or from another World Championship meet in 2019 or 2023.

@@

\jy{This two paragraphs are repetitive. Condense.}
\of{Resolved but needs to be proof read}
To investigate whether RTs at the 2022 World Championships
were significantly different from other competitions, we
used data from male athletes in the 110-meter hurdles and 100-meter
dash, and female athletes in the 100-meter hurdles and 100-meter dash,
provided they competed in the 2022 World Championships and at least
one other competition (2022 national championships, 2019 World
Championships, or 2023 World Championships). Statistical tests showed no
significant differences in RTs between these events, supporting their
inclusion in a unified analysis. However, we excluded data from
200-meter dashes and longer events, as their RTs were found
to be significantly different. This difference is not unexpected, as
RTs are generally less critical in longer sprint events
compared to shorter distances where explosive starts are paramount.
Negative RTs were excluded from the analysis, while positive disqualified RTs
were included due to their low frequency and not being obvious outliers.

  % A
% sensitivity analysis regarding the inclusion of positive disqualified
% RTs is provided in the Supplementary Material.

@@

\jy{Opening sentence not accurate. This is the data to test the first
  hypothesis: national versus World Championship 2022.}
\of{Is it better now?}

%To assess the abnormality of the 2022 World Championships, we examined
%ata beyond World Championship events. This decision was motivated by

To determine if athletes reacted differently at the 2022 World Championships,
our first places that we look are 2022 national championships.
Prior to a formal analysis, we examined how United States (US) athletes
performed at the 2022 US Track and Field Championships, held from June 23--26,
2022, at Hayward Field in Eugene, Oregon. Since this venue also hosted
the 2022 World Championships in August, it
provided a unique opportunity to compare. There were four US
110-meter hurdle athletes—Trey Cunningham, Daniel Roberts,
Grant Holloway, and Devon Allen—who competed in both events. All four
athletes recorded faster RTs in every World Championships race
compared to their performances at the national-level event. Similarly,
all four US 100-meter dash athletes---Marvin Bracy, Fred Kerley,
Travyon Bromwell, and Christian Coleman---also recorded slower RTs at
the US Championships than at the World Championships.
To expand the dataset, we included RTs from 100-meter hurdle
and 100-meter dash athletes who competed in other national
competitions held between May and July 2022 across various
countries. Compiling this data presented challenges, as RTs were not
centrally archived and often required searching country-specific
websites, with many results recorded in native languages.

@@

\jy{Improve the topic sentence: this data is for the second
  hypothesis: 2019 vs 2022; 2023 vs 2022.}
\of{Is this more along the lines of what you are looking for}

This data was compiled to test our second focus, athletes who competed at
the 2022 World Championships and the 2019 or 2023 World Championships. 
In the 2019-2022 comparison,
RTs from 2022 were treated as the `treatment' group, with
2019 serving as the `control' group. Similarly, in the 2022-2023
comparison, RTs from 2022 were treated as the `treatment'
group, with 2023 serving as the `control' group. This structure allowed
us to prepare datasets suitable for examining RTs of athletes
who competed across multiple World Championships.

@@

\jy{write a quick overview paragraph here.}
\of{Is it too much of a re-telling of the introduction?}

In addition to the work presented in Section~\ref{sec:problem1} we thought that
it was relevant and important to expand on prior work that focused on an
appropriate reaction time barrier.  Implemented by World Athletics to discourage
false starts, athletes with RTs below 0.1 seconds are disqualified.  In this
section we will attempt to model historical data to recommend a reaction barrier
based in statistical analysis.
